---
title: "Create Nasal Modules and Gene Set Expression Values"
author: "Elizabeth Whalen"
date: "4/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Optionally install, and load required libraries
```{r loadLibraries}

# Determine missing Bioconductor packages and install
BioCPackages <- c("edgeR", "biomaRt")
new.packages <- BioCPackages[!(BioCPackages %in% installed.packages()[,"Package"])]
if(length(new.packages)) {
  source("http://bioconductor.org/biocLite.R")
  biocLite(new.packages)
}
 
# Determine missing CRAN Packages and install
CRANPackages <- c("WGCNA")
new.packages <- CRANPackages[!(CRANPackages %in% installed.packages()[,"Package"])]
if(length(new.packages)) {
  install.packages(new.packages)
}

# Load libraries further down...

```

## Read in Nasal GEO data

```{r readData}

dataDir <- "../data"
resultsDir <- "../results"


nasalGEO<-read.delim(file=file.path(dataDir, "raw_counts_nasal_muppits523.txt"), sep=",", row.names=1)
dim(nasalGEO)
# get the flow cell info
table(unlist(lapply(strsplit(colnames(nasalGEO), "_"), function(x) {x[[2]]})))

# read in meta data
nasalData<-read.csv(file=file.path(dataDir, "MUPPITS1_update3_nasal_metadata_GEO.csv"), skip=43)
# subset to actual data
nasalData<-nasalData[1:523, 1:44] # keep rows for 523 libraries and keep columns up until raw.file

all(colnames(nasalGEO)==nasalData$Sample.Name)
# need to rearrange
nasalData<-nasalData[match(colnames(nasalGEO), nasalData$Sample.Name),]
all(colnames(nasalGEO)==nasalData$Sample.Name)

```

## Create Voom Counts for Phase 1

```{r phase1Voom}

# subset to phase 1 libraries
keepPhase1<-which(nasalData$Characteristics..Flowcell=="C893JANXX")

nasalCounts1<-nasalGEO[,keepPhase1]
nasalData1<-nasalData[keepPhase1,]
all(colnames(nasalCounts1)==nasalData1$Sample.Name)

library(edgeR)

d1<-DGEList(counts=nasalCounts1)
d1<-calcNormFactors(d1)

keepRows1<-rowSums(round(cpm(d1$counts)) >= 1) >= 0.1*ncol(nasalCounts1)
table(keepRows1)	# removes 44,931 rows

curDGENasal1<-d1[keepRows1,]	# now have 19,322 rows
curDGENasal1<-calcNormFactors(curDGENasal1)

library(biomaRt)

mart = useMart(host='oct2016.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBio1<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=rownames(curDGENasal1), mart=mart)
resultsBioProtein1<-resultsBio1[which(resultsBio1$gene_biotype=="protein_coding"),]
keepIndex1<-which(rownames(curDGENasal1) %in% resultsBioProtein1$ensembl_gene_id)
length(keepIndex1) # 14,057 

curDGENasal1<-curDGENasal1[keepIndex1,]
dim(curDGENasal1)	# 14057    41
curDGENasal1<-calcNormFactors(curDGENasal1)

### Run Voom

# make sure we have factors set up correctly
# for phase 1 my previous DGE object matches this one exactly (counts and norm.factors)
nasalData1$Visit<-as.character(nasalData1$Characteristics..Analysis.Visit)
nasalData1$Visit[which(nasalData1$Visit %in% c("Visit 1a","Visit 2a"))]<-"Visit A"
nasalData1$Visit[which(nasalData1$Visit %in% c("Visit 1b","Visit 2b"))]<-"Visit B"
table(nasalData1$Visit)
nasalData1$Visit<-factor(nasalData1$Visit, levels=c("Visit 0","Visit A","Visit B"))

curMM<-model.matrix(~Visit, data=nasalData1)
nasalVoom1<-voomWithQualityWeights(curDGENasal1, curMM, plot=TRUE)

# now need to take into account duplicate correlation
nasalData1$Source.name<-as.factor(as.character(nasalData1$Source.name))
corfit1<-duplicateCorrelation(nasalVoom1, curMM, block=nasalData1$Source.name)
corfit1$consensus.correlation	# 0.2562075

# matches the values from the original processing
newNasalVoom1<-voomWithQualityWeights(curDGENasal1, curMM, plot=TRUE, block=nasalData1$Patient.ID, correlation=corfit1$consensus)

```

## Create Voom Counts for Phase 2

```{r phase2Voom}

# subset to phase 2 libraries
keepPhase2<-which(nasalData$Characteristics..Flowcell=="C95GDANXX")

nasalCounts2<-nasalGEO[,keepPhase2]
nasalData2<-nasalData[keepPhase2,]
all(colnames(nasalCounts2)==nasalData2$Sample.Name)

# libraries that were not included in the first round of module creation because of more stringent QC
#libsToRemove2<-c("lib12362_C95GDANXX", "lib12399_C95GDANXX", "lib12401_C95GDANXX", "lib12426_C95GDANXX", "lib12427_C95GDANXX", "lib12528_C95GDANXX",
#                 "lib12364_C95GDANXX","lib12380_C95GDANXX","lib12385_C95GDANXX","lib12489_C95GDANXX")
libsToRemove2<-as.character(nasalData2$Sample.Name[which(nasalData2$Characteristics..medianCVCoverage > 1 | nasalData2$Characteristics.percentAligned < 0.78)])

remPhase2<-which(nasalData2$Sample.Name %in% libsToRemove2)
nasalData2<-nasalData2[-remPhase2,]
nasalCounts2<-nasalCounts2[,-remPhase2]
all(colnames(nasalCounts2)==nasalData2$Sample.Name)

d2<-DGEList(counts=nasalCounts2)
d2<-calcNormFactors(d2)

keepRows2<-rowSums(round(cpm(d2$counts)) >= 1) >= 0.1*ncol(nasalCounts2)
table(keepRows2)	# removes 45,776 rows

curDGENasal2<-d2[keepRows2,]	# now have 18,477 rows

mart = useMart(host='oct2016.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBio2<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=rownames(curDGENasal2), mart=mart)
resultsBioProtein2<-resultsBio2[which(resultsBio2$gene_biotype=="protein_coding"),]
keepIndex2<-which(rownames(curDGENasal2) %in% resultsBioProtein2$ensembl_gene_id)
length(keepIndex2) # 13,928 (same as before)

curDGENasal2<-curDGENasal2[keepIndex2,]
dim(curDGENasal2)	# 13928   142

### Run Voom

# make sure we have factors set up correctly
nasalData2$Visit<-as.character(nasalData2$Characteristics..Analysis.Visit)
nasalData2$Visit[which(nasalData2$Visit=="Visit 0 - Screening and Enrollment")]<-"Visit 0"
nasalData2$Visit[which(nasalData2$Visit %in% c("Visit 1a","Visit 2a"))]<-"Visit A"
nasalData2$Visit[which(nasalData2$Visit %in% c("Visit 1b","Visit 2b"))]<-"Visit B"
table(nasalData2$Visit)
nasalData2$Visit<-factor(nasalData2$Visit, levels=c("Visit 0","Visit A","Visit B"))

curMM2<-model.matrix(~Visit, data=nasalData2)
nasalVoom2<-voomWithQualityWeights(curDGENasal2, curMM2, plot=TRUE)

# now need to take into account duplicate correlation
nasalData2$Source.name<-as.factor(as.character(nasalData2$Source.name))
corfit2<-duplicateCorrelation(nasalVoom2, curMM2, block=nasalData2$Source.name)
corfit2$consensus.correlation	# 0.339899 (0.340055)

newNasalVoom2<-voomWithQualityWeights(curDGENasal2, curMM2, plot=TRUE, block=nasalData2$Source.name, correlation=corfit2$consensus)

```

## Create Voom Counts for Phase 3

```{r phase3Voom}

# subset to phase 3 libraries
keepPhase3<-which(nasalData$Characteristics..Flowcell=="CAJMRANXX")

nasalCounts3<-nasalGEO[,keepPhase3]
nasalData3<-nasalData[keepPhase3,]
all(colnames(nasalCounts3)==nasalData3$Sample.Name)

d3<-DGEList(counts=nasalCounts3)
d3<-calcNormFactors(d3)

keepRows3<-rowSums(round(cpm(d3$counts)) >= 1) >= 0.1*ncol(nasalCounts3)
table(keepRows3)	# removes 45,076 rows

curDGENasal3<-d3[keepRows3,]	# now have 19,177 rows

mart3 = useMart(host='dec2016.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBio3<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=rownames(curDGENasal3), mart=mart3)

keepGeneIDs3<-as.character(resultsBio3$ensembl_gene_id[which(resultsBio3$gene_biotype=="protein_coding")])
matchIndex3<-match(keepGeneIDs3, rownames(curDGENasal3))

curDGENasal3<-curDGENasal3[matchIndex3,]
dim(curDGENasal3)  # now have 14,184 rows

### Run Voom

# make sure we have factors set up correctly
nasalData3$Visit<-as.character(nasalData3$Characteristics..Analysis.Visit)
nasalData3$Visit[which(nasalData3$Visit=="Visit 0 - Screening and Enrollment")]<-"Visit 0"
nasalData3$Visit[which(nasalData3$Visit %in% c("Visit 1a","Visit 2a"))]<-"Visit A"
nasalData3$Visit[which(nasalData3$Visit %in% c("Visit 1b","Visit 2b"))]<-"Visit B"
table(nasalData3$Visit)
nasalData3$Visit<-factor(nasalData3$Visit, levels=c("Visit 0","Visit A","Visit B"))

curMM3<-model.matrix(~Visit, data=nasalData3)

nasalVoom3<-voomWithQualityWeights(curDGENasal3, curMM3, plot=TRUE)

# now need to take into account duplicate correlation
nasalData3$Source.name<-as.factor(as.character(nasalData3$Source.name))
corfit3<-duplicateCorrelation(nasalVoom3, curMM3, block=nasalData3$Source.name)
corfit3$consensus.correlation	# 0.2737985 (0.2738104)

newNasalVoom3<-voomWithQualityWeights(curDGENasal3, curMM3, plot=TRUE, block=nasalData3$Source.name, correlation=corfit3$consensus)

```

## Create Voom Counts for Phase 4

```{r phase4Voom}

# subset to phase 4 libraries
keepPhase4<-which(nasalData$Characteristics..Flowcell %in% c("CB4FNANXX","CB5HFANXX"))

nasalCounts4<-nasalGEO[,keepPhase4]
nasalData4<-nasalData[keepPhase4,]
all(colnames(nasalCounts4)==nasalData4$Sample.Name)

d4<-DGEList(counts=nasalCounts4)
d4<-calcNormFactors(d4)

# filter genes/rows
keepRows4<-rowSums(round(cpm(d4$counts)) >= 1) >= 0.1*ncol(nasalCounts4)
table(keepRows4) # removes 45,737 rows

curDGENasal4<-d4[keepRows4,]	# now have 18,516 rows

mart4 = useMart(host='may2017.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBio4<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=rownames(curDGENasal4), mart=mart4)

keepGeneIDs4<-as.character(resultsBio4$ensembl_gene_id[which(resultsBio4$gene_biotype=="protein_coding")])
matchIndex4<-match(keepGeneIDs4, rownames(curDGENasal4))

curDGENasal4<-curDGENasal4[matchIndex4,]
dim(curDGENasal4)  # now have 13,960 rows

### Run Voom

# make sure we have factors set up correctly
nasalData4$Visit<-as.character(nasalData4$Characteristics..Analysis.Visit)
nasalData4$Visit[which(nasalData4$Visit=="Visit 0 - Screening and Enrollment")]<-"Visit 0"
nasalData4$Visit[which(nasalData4$Visit %in% c("Visit 1a","Visit 2a"))]<-"Visit A"
nasalData4$Visit[which(nasalData4$Visit %in% c("Visit 1b","Visit 2b"))]<-"Visit B"
table(nasalData4$Visit)
nasalData4$Visit<-factor(nasalData4$Visit, levels=c("Visit 0","Visit A","Visit B"))

curMM4<-model.matrix(~Visit, data=nasalData4)
nasalVoom4<-voomWithQualityWeights(curDGENasal4, curMM4, plot=TRUE)

# now need to take into account duplicate correlation
nasalData4$Source.name<-as.factor(as.character(nasalData4$Source.name))
corfit4<-duplicateCorrelation(nasalVoom4, curMM4, block=nasalData4$Source.name)
corfit4$consensus.correlation	# 0.2538339

newNasalVoom4<-voomWithQualityWeights(curDGENasal4, curMM4, plot=TRUE, block=nasalData4$Source.name, correlation=corfit4$consensus)

```

## Subset to Libraries Used in Module Creation

```{r subsetForModules}

# Combine Voom Counts
newNasalVoom1sub<-newNasalVoom1[which(rownames(newNasalVoom1) %in% rownames(newNasalVoom2)),]
newNasalVoom2sub<-newNasalVoom2[which(rownames(newNasalVoom2) %in% rownames(newNasalVoom1)),]
dim(newNasalVoom1sub)
dim(newNasalVoom2sub)
sum(rownames(newNasalVoom1sub) %in% rownames(newNasalVoom2sub)) # make sure the rows are ordered the same before combining
newNasalVoom1sub<-newNasalVoom1sub[match(rownames(newNasalVoom2sub), rownames(newNasalVoom1sub)),]
all(rownames(newNasalVoom1sub) == rownames(newNasalVoom2sub))

modNasalVoom<-cbind(newNasalVoom1sub, newNasalVoom2sub)
# subset modNasalVoom to the 146 libs used in module construction
nasalDataModule<-nasalData[which(nasalData$Characteristics..Used.in.Module.Construction=="Yes"),]
keepLibs<-which(colnames(modNasalVoom) %in% nasalDataModule$Sample.Name)
modNasalVoomOrig<-modNasalVoom
modNasalVoom<-modNasalVoomOrig[,keepLibs]

all(colnames(modNasalVoom)==nasalDataModule$Sample.Name)

modNDesign<-nasalDataModule
all(colnames(modNasalVoom)==modNDesign$Sample.Name)

```

## Reset Cell Percentages

When the modules were created, we did not have cell percentages for visit B so visit B percentages were set to visit A percentages.

```{r resetCellPerc}

neut<-modNDesign$Characteristics..Nasal.Neutrophil.Percentage 
lymp<-modNDesign$Characteristics..Nasal.Lymphocyte.Percentage 
eos<-modNDesign$Characteristics..Nasal.Eosinophil.Percentage 
mac<-modNDesign$Characteristics..Nasal.Macrophage.Percentage 
epi<-modNDesign$Characteristics..Nasal.Epithelial.Percentage
squa<-modNDesign$Characteristics..Nasal.Squamous.Percentage

uniIndivs<-unique(as.character(modNDesign$Source.name))
for (i in 1:length(uniIndivs))
{
  matchIndex<-which(uniIndivs[i]==modNDesign$Source.name)
  matchVisit<-as.character(modNDesign$Characteristics..Analysis.Visit[matchIndex])
  if (length(grep("b", matchVisit)) > 0)
  {
    # find out which visit it is
    visitNum<-substr(matchVisit[grep("b", matchVisit)], 7, 7)
    if (length(visitNum)==2)
    {
      # have both visits 1 and 2
      if (any(matchVisit=="Visit 1a"))
      {
        # only get the a visit from the full set
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 1b")
        aIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 1a")
        neut[bIndex]<-neut[aIndex]
        lymp[bIndex]<-lymp[aIndex]
        eos[bIndex]<-eos[aIndex]
        mac[bIndex]<-mac[aIndex]
        epi[bIndex]<-epi[aIndex]
        squa[bIndex]<-squa[aIndex]
      }
      if (!("Visit 1a" %in% matchVisit))
      {
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 1b")
        neut[bIndex]<-NA
        lymp[bIndex]<-NA
        eos[bIndex]<-NA
        mac[bIndex]<-NA
        epi[bIndex]<-NA
        squa[bIndex]<-NA
      }
      
      if (any(matchVisit=="Visit 2a"))
      {
        # only get the a visit from the full set
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 2b")
        aIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 2a")
        neut[bIndex]<-neut[aIndex]
        lymp[bIndex]<-lymp[aIndex]
        eos[bIndex]<-eos[aIndex]
        mac[bIndex]<-mac[aIndex]
        epi[bIndex]<-epi[aIndex]
        squa[bIndex]<-squa[aIndex]
      }
      if (!("Visit 2a" %in% matchVisit))
      {
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit=="Visit 2b")
        neut[bIndex]<-NA
        lymp[bIndex]<-NA
        eos[bIndex]<-NA
        mac[bIndex]<-NA
        epi[bIndex]<-NA
        squa[bIndex]<-NA
      }
    }
    else
    {
      otherVisit<-paste("Visit ",visitNum,"a", sep="")
      if (any(matchVisit==otherVisit))
      {
        # only get the a visit from the full set
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit==paste("Visit ",visitNum,"b", sep=""))
        aIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit==paste("Visit ",visitNum,"a", sep=""))
        neut[bIndex]<-neut[aIndex]
        lymp[bIndex]<-lymp[aIndex]
        eos[bIndex]<-eos[aIndex]
        mac[bIndex]<-mac[aIndex]
        epi[bIndex]<-epi[aIndex]
        squa[bIndex]<-squa[aIndex]
      }
      else
      {
        # there's no matching visit A for this visit B so need to set to NA
        bIndex<-which(uniIndivs[i]==modNDesign$Source.name & modNDesign$Characteristics..Analysis.Visit==paste("Visit ",visitNum,"b", sep=""))
        neut[bIndex]<-NA
        lymp[bIndex]<-NA
        eos[bIndex]<-NA
        mac[bIndex]<-NA
        epi[bIndex]<-NA
        squa[bIndex]<-NA
      }
    }
  }
}
# there were a few visit A values for visit B that we had back in 2016 so will use those from the full set
# and some were missing so need to rearrange some values

# had to hard code these in
neut[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12502", "lib12503", "lib12524", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib12366", "lib12416", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  neut[curIndex]<-nasalData$Characteristics..Nasal.Neutrophil.Percentage[matchIndex[1]]
}

lymp[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12502", "lib12503", "lib12524", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib12366", "lib12416", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  lymp[curIndex]<-nasalData$Characteristics..Nasal.Lymphocyte.Percentage[matchIndex[1]]
}

eos[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12502", "lib12503", "lib12524", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib12366", "lib12416", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  eos[curIndex]<-nasalData$Characteristics..Nasal.Eosinophil.Percentage[matchIndex[1]]
}

mac[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12502", "lib12503", "lib12524", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib12366", "lib12416", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  mac[curIndex]<-nasalData$Characteristics..Nasal.Macrophage.Percentage[matchIndex[1]]
}

epi[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib11122", "lib12366", "lib12416", "lib12428", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  epi[curIndex]<-nasalData$Characteristics..Nasal.Epithelial.Percentage[matchIndex[1]]
}

squa[which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib12437", "lib12438", "lib12535", "lib12536"))]<-NA
setToNumberIndex<-which(unlist(lapply(strsplit(as.character(modNDesign$Sample.Name),"_"), function(x) {x[[1]]})) %in% c("lib11094", "lib11122", "lib12366", "lib12416", "lib12428", "lib12456")) # these are all visit 1b
for (i in 1:length(setToNumberIndex))
{
  curIndex<-setToNumberIndex[i]
  curIndiv<-as.character(modNDesign$Source.name[curIndex])
  matchIndex<-which(curIndiv==nasalData$Source.name & nasalData$Characteristics..Analysis.Visit=="Visit 1a")
  squa[curIndex]<-nasalData$Characteristics..Nasal.Squamous.Percentage[matchIndex[1]]
}

```

## Assign to Cell Type

```{r assignCellType}

neutP<-c()
neutCor<-c()
lympP<-c()
lympCor<-c()
eosP<-c()
eosCor<-c()
macP<-c()
macCor<-c()
epiP<-c()
epiCor<-c()
squaP<-c()
squaCor<-c()
for (i in 1:nrow(modNasalVoom))
{
  corVals<-cor.test(modNasalVoom$E[i,], neut, use="pairwise.complete.obs")
  neutP<-c(neutP, corVals$p.value)
  neutCor<-c(neutCor, corVals$estimate)
  
  corVals<-cor.test(modNasalVoom$E[i,], lymp, use="pairwise.complete.obs")
  lympP<-c(lympP, corVals$p.value)
  lympCor<-c(lympCor, corVals$estimate)

  corVals<-cor.test(modNasalVoom$E[i,], eos, use="pairwise.complete.obs")
  eosP<-c(eosP, corVals$p.value)
  eosCor<-c(eosCor, corVals$estimate)
  
  corVals<-cor.test(modNasalVoom$E[i,], mac, use="pairwise.complete.obs")
  macP<-c(macP, corVals$p.value)
  macCor<-c(macCor, corVals$estimate)
  
  corVals<-cor.test(modNasalVoom$E[i,], epi, use="pairwise.complete.obs")
  epiP<-c(epiP, corVals$p.value)
  epiCor<-c(epiCor, corVals$estimate)
  
  corVals<-cor.test(modNasalVoom$E[i,], squa, use="pairwise.complete.obs")
  squaP<-c(squaP, corVals$p.value)
  squaCor<-c(squaCor, corVals$estimate)
}

# adjust p-values
neutFDR<-p.adjust(neutP)
lympFDR<-p.adjust(lympP)
eosFDR<-p.adjust(eosP)
macFDR<-p.adjust(macP)
epiFDR<-p.adjust(epiP)
squaFDR<-p.adjust(squaP)

# set the cutoffs for assignment to cell type
# exact match
sum(neutFDR < 0.05 & neutCor > 0) # 1525
sum(lympFDR < 0.05 & lympCor > 0) # 1192
sum(eosFDR < 0.05 & eosCor > 0) # 672
sum(macFDR < 0.05 & macCor > 0) # 788
sum(epiFDR < 0.05 & epiCor > 0) # 241
sum(squaFDR < 0.05 & squaCor > 0) # 1058

neutGenes<-rownames(modNasalVoom)[which(neutFDR < 0.05 & neutCor > 0)]
lympGenes<-rownames(modNasalVoom)[which(lympFDR < 0.05 & lympCor > 0)]
eosGenes<-rownames(modNasalVoom)[which(eosFDR < 0.05 & eosCor > 0)]
macGenes<-rownames(modNasalVoom)[which(macFDR < 0.05 & macCor > 0)]
epiGenes<-rownames(modNasalVoom)[which(epiFDR < 0.05 & epiCor > 0)]
squaGenes<-rownames(modNasalVoom)[which(squaFDR < 0.05 & squaCor > 0)]
unmaGenes<-rownames(modNasalVoom)[which(!(rownames(modNasalVoom) %in% unique(c(neutGenes, lympGenes, eosGenes, macGenes, epiGenes, squaGenes))))]

```

## Now Run WGCNA and Create Modules

```{r runWGCNA}

library(WGCNA)

# subset to cell specific genes
#matchIndex<-match(neutGenes, rownames(modNasalVoom))
#matchIndex<-match(lympGenes, rownames(modNasalVoom))
#matchIndex<-match(eosGenes, rownames(modNasalVoom))
#matchIndex<-match(macGenes, rownames(modNasalVoom))
#matchIndex<-match(epiGenes, rownames(modNasalVoom))
#matchIndex<-match(squaGenes, rownames(modNasalVoom))
#matchIndex<-match(unmaGenes, rownames(modNasalVoom))

#setVoom<-modNasalVoom[matchIndex,]
# power for
# neut - 12
# lymp - 12
# eos - 8
# mac - 12
# epi - 12
# squa - 12
# unassigned - 10

#powers<-c(1:10,seq(12,30,2))
# call the network topology analysis function
#sft<-pickSoftThreshold(t(setVoom$E), powerVector=powers, verbose=5)
# plot the results
#par(mfrow=c(1,2))
# scale-free topology fit index as a function of the soft-thresholding power
#plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab="Soft Threshold (power)", ylab="Scale Free Topology Model Fit,signed R^2", type="n", main=paste("Scale independence"))
#text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels=powers, col="red")
# this line corresponds to using a R^2 cutoff of h
#abline(h=0.9, col="red")
# mean connectivity as a function of the soft-thresholding power
#plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab="Soft Threshold (power)", ylab="Mean Connectivity", type="n", main=paste("Mean Connectivity"))
#text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, col="red")



createWGCNA<-function(cellGenes, allNasalVoom, curPower)
{
	matchIndex<-match(cellGenes, rownames(allNasalVoom))
	setVoom<-allNasalVoom[matchIndex,]
	setVoom<-setVoom[order(rownames(setVoom)),]
	
	time1<-Sys.time()
	voomNetNasal<-blockwiseModules(t(setVoom$E), power=curPower, networkType="signed", TOMType="signed", minModuleSize=30, maxBlockSize=6000, saveTOMS=TRUE, saveTOMFileBase = "test", deepSplit=4, numericLabels=TRUE, minCoreKME=0.7, minKMEtoStay=0.5)
	time2<-Sys.time()
	print(time2-time1)	# 10 sec, 7 sec, 4 sec, 5 sec, 3 sec, 7 sec, 42 sec
	return(voomNetNasal)
}
# 6 sets for neut, 4 lymp, 5 eos, 4 mac, 1 epi, 1 squa, 31 for unmatched
neutrophilWGCNA<-createWGCNA(neutGenes, modNasalVoom, curPower=12)
lymphocyteWGCNA<-createWGCNA(lympGenes, modNasalVoom, curPower=12)
eosinophilWGCNA<-createWGCNA(eosGenes, modNasalVoom, curPower=8)
macrophageWGCNA<-createWGCNA(macGenes, modNasalVoom, curPower=12)
epithelialWGCNA<-createWGCNA(epiGenes, modNasalVoom, curPower=12)
squamousWGCNA<-createWGCNA(squaGenes, modNasalVoom, curPower=12)
unmatchedWGCNA<-createWGCNA(unmaGenes, modNasalVoom, curPower=10)

# now how to store it
# need to order the gene names because that was done in the function createWGCNA
neutSet<-cbind(rep("neutrophil", length(neutGenes)), neutGenes[order(neutGenes)], paste("neut",neutrophilWGCNA$colors,sep=""))
lympSet<-cbind(rep("lymphocyte", length(lympGenes)), lympGenes[order(lympGenes)], paste("lymp",lymphocyteWGCNA$colors,sep=""))
eosSet<-cbind(rep("eosinophil", length(eosGenes)), eosGenes[order(eosGenes)], paste("eos",eosinophilWGCNA$colors,sep=""))
macSet<-cbind(rep("macrophage", length(macGenes)), macGenes[order(macGenes)], paste("mac",macrophageWGCNA$colors,sep=""))
epiSet<-cbind(rep("epithelial", length(epiGenes)), epiGenes[order(epiGenes)], paste("epi",epithelialWGCNA$colors,sep=""))
squaSet<-cbind(rep("squamous", length(squaGenes)), squaGenes[order(squaGenes)], paste("squa",squamousWGCNA$colors,sep=""))
unmaSet<-cbind(rep("unassigned", length(unmaGenes)), unmaGenes[order(unmaGenes)], paste("unma",unmatchedWGCNA$colors,sep=""))

allSets<-rbind(neutSet, lympSet, eosSet, macSet, epiSet, squaSet, unmaSet)
table(allSets[,3])

colnames(allSets)<-c("cellType","ensemblID","geneSet")
allSets<-as.data.frame(allSets)
# now add gene symbols
allSets$symbol<-""
for (i in 1:nrow(allSets))
{
  curID<-as.character(allSets$ensemblID[i])
  allSets$symbol[i]<-as.character(resultsBioProtein1$hgnc_symbol[which(curID==resultsBioProtein1$ensembl_gene_id)  ])
}

# use a different mart to set gene symbol to make it match
mart = useMart(host='dec2016.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBioMod<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=allSets$ensemblID, mart=mart)

# now add gene symbols
allSets$symbol<-""
for (i in 1:nrow(allSets))
{
  curID<-as.character(allSets$ensemblID[i])
  allSets$symbol[i]<-as.character(resultsBioMod$hgnc_symbol[which(curID==resultsBioMod$ensembl_gene_id)  ])
}

# Write out the nasal modules as a .csv file.
write.csv(allSets[allSets$cellType == "neutrophil", ], file=file.path(resultsDir, "neutrophilNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "lymphocyte", ], file=file.path(resultsDir, "lymphocyteNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "eosinophil", ], file=file.path(resultsDir, "eosinophilNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "macrophage", ], file=file.path(resultsDir, "macrophageNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "epithelial", ], file=file.path(resultsDir, "epithelialNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "squamous",   ], file=file.path(resultsDir, "squamousNasalModules.csv"),   quote=F, row.names=F)
write.csv(allSets[allSets$cellType == "unassigned", ], file=file.path(resultsDir, "unassignedNasalModules.csv"), quote=F, row.names=F)
write.csv(allSets, file=file.path(resultsDir, "allNasalModules.csv"), quote=F, row.names=F)

```

## Redo Voom Counts for Phase 2 with New QC Cutoffs

```{r phase2VoomRedo}

# subset to phase 2 libraries
keepPhase2<-which(nasalData$Characteristics..Flowcell=="C95GDANXX")

nasalCounts2<-nasalGEO[,keepPhase2]
nasalData2<-nasalData[keepPhase2,]
all(colnames(nasalCounts2)==nasalData2$Sample.Name)

d2<-DGEList(counts=nasalCounts2)
d2<-calcNormFactors(d2)

keepRows2<-rowSums(round(cpm(d2$counts)) >= 1) >= 0.1*ncol(nasalCounts2)
table(keepRows2)	# removes 45,709 rows

curDGENasal2<-d2[keepRows2,]	# now have 18,544 rows

mart = useMart(host='dec2016.archive.ensembl.org', biomart='ENSEMBL_MART_ENSEMBL', dataset="hsapiens_gene_ensembl")
resultsBio2<-getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters="ensembl_gene_id", values=rownames(curDGENasal2), mart=mart)

keepGeneIDs2<-as.character(resultsBio2$ensembl_gene_id[which(resultsBio2$gene_biotype=="protein_coding")])
matchIndex2<-match(keepGeneIDs2, rownames(curDGENasal2))

curDGENasal2<-curDGENasal2[matchIndex2,]
dim(curDGENasal2)	# 13955   152

### Run Voom

# make sure we have factors set up correctly
nasalData2$Visit<-as.character(nasalData2$Characteristics..Analysis.Visit)
nasalData2$Visit[which(nasalData2$Visit=="Visit 0 - Screening and Enrollment")]<-"Visit 0"
nasalData2$Visit[which(nasalData2$Visit %in% c("Visit 1a","Visit 2a"))]<-"Visit A"
nasalData2$Visit[which(nasalData2$Visit %in% c("Visit 1b","Visit 2b"))]<-"Visit B"
table(nasalData2$Visit)
nasalData2$Visit<-factor(nasalData2$Visit, levels=c("Visit 0","Visit A","Visit B"))

curMM2<-model.matrix(~Visit, data=nasalData2)
nasalVoom2<-voomWithQualityWeights(curDGENasal2, curMM2, plot=TRUE)

# now need to take into account duplicate correlation
nasalData2$Source.name<-as.factor(as.character(nasalData2$Source.name))
corfit2<-duplicateCorrelation(nasalVoom2, curMM2, block=nasalData2$Source.name)
corfit2$consensus.correlation	# 0.3105927 (0.3132524)

newNasalVoom2<-voomWithQualityWeights(curDGENasal2, curMM2, plot=TRUE, block=nasalData2$Source.name, correlation=corfit2$consensus)

```

## Combine All Voom Counts

```{r combineAllVoom}

# find common genes
allRN<-c(rownames(newNasalVoom1), rownames(newNasalVoom2), rownames(newNasalVoom3), rownames(newNasalVoom4))
keepGenes<-names(which(table(allRN)>=4))
length(keepGenes) # 13,672

# subset voom counts to just genes in common
nasalVoom1Sub<-newNasalVoom1[match(keepGenes, rownames(newNasalVoom1)),]
nasalVoom2Sub<-newNasalVoom2[match(keepGenes, rownames(newNasalVoom2)),]
nasalVoom3Sub<-newNasalVoom3[match(keepGenes, rownames(newNasalVoom3)),]
nasalVoom4Sub<-newNasalVoom4[match(keepGenes, rownames(newNasalVoom4)),]
all(rownames(nasalVoom1Sub)==rownames(nasalVoom2Sub))
all(rownames(nasalVoom1Sub)==rownames(nasalVoom3Sub))
all(rownames(nasalVoom1Sub)==rownames(nasalVoom4Sub))

# make the design matrices compatible - all design matrices have 3 columns so it's okay
allNasalVoom<-cbind(nasalVoom1Sub, nasalVoom2Sub, nasalVoom3Sub, nasalVoom4Sub)
save(allNasalVoom, file=file.path(resultsDir, "allNasalVoom.RData"))
```

## Remove Samples that are not in Final Set for Analysis

```{r removeSamples}

keepSamples<-as.character(nasalData$Sample.Name[which(nasalData$Characteristics..Used.in.Analaysis=="Yes")])
matchIndex<-match(keepSamples, colnames(allNasalVoom))

allNasalVoomSub<-allNasalVoom[, matchIndex]


```

## Calculate Gene Set Mean Expression

```{r geneSetExpression}

uniMods<-as.character(unique(allSets$geneSet))
gsExpr<-c()
for (i in 1:length(uniMods))
{
  curMod<-uniMods[i]
  curIDs<-as.character(allSets$ensemblID[which(allSets$geneSet==curMod)])
  matchIndex<-match(curIDs, rownames(allNasalVoomSub))
  if (any(is.na(matchIndex)))
    matchIndex<-matchIndex[-which(is.na(matchIndex))]
  if (length(matchIndex) > 1)
    gsExpr<-rbind(gsExpr, apply(allNasalVoomSub$E[matchIndex,], 2, mean))  
  else
    gsExpr<-rbind(gsExpr, allNasalVoomSub$E[matchIndex,])
}
rownames(gsExpr)<-uniMods
nasalGSExpr <- gsExpr
save(nasalGSExpr, file=file.path(resultsDir, paste("Nasal", "gsExpr.RData", sep="_")))

```


