---
title: "MUPPITS Tempus Mean/SE Plots Module Level Visits A&B"
author: "Elizabeth Whalen"
date: "11/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in Data

```{r readData}

dataDir <- "../data"
resultsDir <- "../results"

tGS<-read.csv(file=file.path(dataDir, "tempus_Combine4Phases/geneSets_v2_using176samples/tempus_geneSetsByAssignedCell_v2.csv"))

# read in voom, 
tGSmeans<-read.csv(file=file.path(dataDir, "tempusGeneSets_387samples.csv"), row.names=1)

all(rownames(tGSmeans) %in% tGS$geneSet)

tDesign<-read.csv(file=file.path(dataDir, "tempus_design_all4Phases_387samples_update.csv"))

#load(file="/Users/ewhalen/Box Sync/Projects/ICAC/MUPPITS/tempus_Combine4Phases/BestVersion/tempusVoom_387samples.RData")  
#tVoom<-tempusVoomNoRepsv2

#all(colnames(tVoom)==tDesign$LibraryID)
all(colnames(tGSmeans)==tDesign$LibraryID)

```

## Remove Batch Effects

```{r removeBatch}

# remove visit 0 and then remove batch effects
rem0<-which(tDesign$Visit=="Visit 0")
tDesignAB<-tDesign[-rem0,]
tGSmeansAB<-tGSmeans[,-rem0]
all(tDesignAB$library.sampleId==colnames(tGSmeansAB))
  
# set up viral type
tDesignAB$Viral.Type.at.Visit2<-factor(as.character(tDesignAB$Viral.Type.at.Visit2), levels=c("Non-viral","Viral"))
table(tDesignAB$Viral.Type.at.Visit2)

# set up cell percentages - they are already numeric and % 
# NOTE: a few do not add up to 100 - should these be changed?
range(apply(tDesignAB[,70:74], 1, sum), na.rm=T)
hist(apply(tDesignAB[,70:74], 1, sum))
which(apply(tDesignAB[,70:74], 1, sum) < 98)
sort(apply(tDesignAB[,70:74], 1, sum))

remIndex<-unique(c(which(is.na(tDesignAB$Blood.Eosinophil.Differential)), which(is.na(tDesignAB$Blood.Lymphocyte.Differential)), which(is.na(tDesignAB$Blood.Monocyte.Differential)), which(is.na(tDesignAB$Blood.Basophil.Differential)), which(is.na(tDesignAB$Viral.Type.at.Visit2)), which(is.na(tDesignAB$Viral.Type.at.Visit2))))
length(remIndex) # lost 65 out of 285 samples
dim(tDesignAB)

# missing mainly visit B
table(tDesignAB$Visit[which(is.na(tDesignAB$Blood.Eosinophil.Differential))])

tDesignAB$Visit<-factor(as.character(tDesignAB$Visit), levels=c("Visit a","Visit b"))
tDesignAB$Case.or.Control.Status.Original<-factor(as.character(tDesignAB$Case.or.Control.Status.Original), levels=c("Control","Case"))

rbeMM<-model.matrix(~ Blood.Eosinophil.Differential + Blood.Lymphocyte.Differential + Blood.Monocyte.Differential + Blood.Basophil.Differential + Visit + libCounts, data=tDesignAB[-remIndex,])
dim(rbeMM)
rbeMM[1:3,]

keepmm<-model.matrix(~Case.or.Control.Status.Original + Viral.Type.at.Visit2, data=tDesignAB[-remIndex,])
dim(keepmm)
keepmm[1:3,]

# remove batch effect at module level
tGSmeansABRBE<-removeBatchEffect(tGSmeansAB[,-remIndex], covariates=rbeMM[,-1], design=keepmm)

dim(tGSmeansABRBE)
tGSmeansABRBE[1:5,1:5]

# also need to subset the tempus design file
tDesignABRBE<-tDesignAB[-remIndex,]
all(colnames(tGSmeansABRBE)==tDesignABRBE$library.sampleId)


# also remove virus type as a batch effect
rbeMMV<-model.matrix(~ Blood.Eosinophil.Differential + Blood.Lymphocyte.Differential + Blood.Monocyte.Differential + Blood.Basophil.Differential + Viral.Type.at.Visit2 + Visit + libCounts, data=tDesignAB[-remIndex,])
dim(rbeMMV)
rbeMMV[1:3,]

keepmmV<-model.matrix(~Case.or.Control.Status.Original, data=tDesignAB[-remIndex,])
dim(keepmmV)
keepmmV[1:3,]

# remove batch effect at voom level
tGSmeansABRBEV<-removeBatchEffect(tGSmeansAB[,-remIndex], covariates=rbeMMV[,-1], design=keepmmV)

dim(tGSmeansABRBEV)
tGSmeansABRBEV[1:5,1:5]

# make sure everything matches
all(colnames(tGSmeansABRBEV)==tDesignABRBE$library.sampleId)

```

## Remove Cases on Steroids

```{r removeCaseOnCS}

remCS<-which(tDesignABRBE$CSteroid.Start.Relative.to.Visit=="Before")
tGSmeansABRBE<-tGSmeansABRBE[,-remCS]
tGSmeansABRBEV<-tGSmeansABRBEV[,-remCS]
tDesignABRBE<-tDesignABRBE[-remCS,]
all(tDesignABRBE$library.sampleId==colnames(tGSmeansABRBE))
all(tDesignABRBE$library.sampleId==colnames(tGSmeansABRBEV))

```

## Remove Zero Modules

```{r removeZeroMods}

remMods<-c("neut0","lymp0","eos0","mono0","unma0")
tGSmeansABRBE<-tGSmeansABRBE[-which(rownames(tGSmeansABRBE) %in% remMods),]
tGSmeansABRBEV<-tGSmeansABRBEV[-which(rownames(tGSmeansABRBEV) %in% remMods),]

```

## Mean/SE plots

### Visits A & B: case vs. control

```{r visitABcaseControl}

table(tDesignABRBE$Case.or.Control.Status.Original)

modNames<-rownames(tGSmeansABRBE)
modNames[17:42]<-paste("m",substr(modNames[17:42], 5, nchar(modNames[17:42])),sep="")
modNames<-paste(modNames,".b",sep="")

for (i in 1:nrow(tGSmeansABRBEV))
{
  curMod<-rownames(tGSmeansABRBEV)[i]
  curFN<-paste(curMod, ".png", sep="")

  statMean<-tapply(tGSmeansABRBEV[i,], tDesignABRBE$Case.or.Control.Status.Original, mean)  
  statSD<-tapply(tGSmeansABRBEV[i,], tDesignABRBE$Case.or.Control.Status.Original, sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBEV[i,], tDesignABRBE$Case.or.Control.Status.Original, length))
  upLim<-2^(statMean + 1.96*statSE)
  lowLim<-2^(statMean - 1.96*statSE)
  
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:2, y=2^statMean, pch=19, col=c(rgb(0,0,0),rgb(1,0,0)), cex=3, xlim=c(0.7,2.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:2, labels=c("Control","Case"))
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(1,0,0,alpha=0.5), border=rgb(1,0,0,alpha=0.5))
  
  points(x=1:2, y=2^statMean, pch=19, col=c(rgb(0,0,0),rgb(1,0,0)), cex=3)
  dev.off()
}

```

### Visits A & B: case vs. control V virus + vs. -

```{r visitABcaVcoVposVneg}

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2)

for (i in 1:nrow(tGSmeansABRBE))
{
  curMod<-rownames(tGSmeansABRBE)[i]
  curFN<-paste(curMod, ".png", sep="")

  statMean<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), mean)  
  statSD<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_virusPosVNeg_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(221/255,160/255,221/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.Non-viral","Case.Non-viral","Control.Viral","Case.Viral"), cex.axis=0.6)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(221/255,160/255,221/255, alpha=0.5), border=rgb(221/255,160/255,221/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(160/255,32/255,240/255,alpha=0.5), border=rgb(160/255,32/255,240/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(221/255,160/255,221/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3)
  dev.off()
}

```


### Visits A & B: case vs. control V virus + vs. - New Colors and plotting characters (5/11/18)

```{r visitABcaVcoVposVneg}

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2)

for (i in 1:nrow(tGSmeansABRBE))
{
  curMod<-rownames(tGSmeansABRBE)[i]
  curFN<-paste(curMod, ".png", sep="")

  statMean<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), mean)  
  statSD<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_virusPosVNeg_newColor_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=c(15,15,19,19), col=c(rgb(84/255,84/255,84/255),rgb(197/255,90/255,17/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.Non-viral","Case.Non-viral","Control.Viral","Case.Viral"), cex.axis=0.6)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(197/255,90/255,17/255, alpha=0.5), border=rgb(197/255,90/255,17/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(160/255,32/255,240/255,alpha=0.5), border=rgb(160/255,32/255,240/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=c(15,15,19,19), col=c(rgb(84/255,84/255,84/255),rgb(197/255,90/255,17/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3)
  dev.off()
}


# 1/14/19 remake as svg for paper
for (i in 1:nrow(tGSmeansABRBE))
{
  curMod<-rownames(tGSmeansABRBE)[i]
  curFN<-paste(curMod, ".svg", sep="")

  statMean<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), mean)  
  statSD<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  svg(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_virusPosVNeg_newColor_svg_"), curFN, sep=""), width=3, height=4, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=c(15,15,19,19), col=c(rgb(84/255,84/255,84/255),rgb(197/255,90/255,17/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.Non-viral","Case.Non-viral","Control.Viral","Case.Viral"), cex.axis=0.6)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(197/255,90/255,17/255, alpha=0.5), border=rgb(197/255,90/255,17/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(160/255,32/255,240/255,alpha=0.5), border=rgb(160/255,32/255,240/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=c(15,15,19,19), col=c(rgb(84/255,84/255,84/255),rgb(197/255,90/255,17/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3)
  dev.off()
}

```


### Visits A & B: case vs. control V virus + vs. - Lower Limit is not Set to Zero

```{r visitABcaVcoVposVnegNot0}

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2)

for (i in 1:nrow(tGSmeansABRBE))
{
  curMod<-rownames(tGSmeansABRBE)[i]
  curFN<-paste(curMod, ".png", sep="")

  statMean<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), mean)  
  statSD<-tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBE[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Viral.Type.at.Visit2), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_virusPosVNeg_LLnotSetTo0_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(221/255,160/255,221/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3, xlim=c(0.7,4.3), ylim=c(min(lowLim), max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.Non-viral","Case.Non-viral","Control.Viral","Case.Viral"), cex.axis=0.6)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(221/255,160/255,221/255, alpha=0.5), border=rgb(221/255,160/255,221/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(160/255,32/255,240/255,alpha=0.5), border=rgb(160/255,32/255,240/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(221/255,160/255,221/255),rgb(0,0,0),rgb(160/255,32/255,240/255)), cex=3)
  dev.off()
}

```

### Visits A & B: case vs. control V bmi 95th percentile yes vs. no

```{r visitABbmi}

# set up BMI percentile cutoff of 95th percentile
hist(tDesignABRBE$BMI.Percentile.at.Visit, breaks=20)
tDesignABRBE$BMIcat95<-"low"
tDesignABRBE$BMIcat95[which(tDesignABRBE$BMI.Percentile.at.Visit > 95)]<-"high"
tDesignABRBE$BMIcat95[which(is.na(tDesignABRBE$BMI.Percentile.at.Visit))]<-NA
table(tDesignABRBE$BMIcat95)
tDesignABRBE$BMIcat95<-factor(tDesignABRBE$BMIcat95, levels=c("low","high"))

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$BMIcat95)

# visits A&B: case vs. control as well as BMI95th percentile
for (i in 1:nrow(tGSmeansABRBEV))
{
  curMod<-rownames(tGSmeansABRBEV)[i]
  curFN<-paste(curMod, ".png", sep="")
  
  statMean<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$BMIcat95), mean) 
  statSD<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$BMIcat95), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$BMIcat95), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  # chose darkolivegreen3 for case BMI low over darkolivegreen1 (too light)
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_BMI95th_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(162/255,205/255,90/255),rgb(0,0,0),rgb(85/255,107/255,47/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.lowBMI","Case.lowBMI","Control.highBMI","Case.highBMI"), cex.axis=0.8)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(162/255,205/255,90/255, alpha=0.5), border=rgb(162/255,205/255,90/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(85/255,107/255,47/255,alpha=0.5), border=rgb(85/255,107/255,47/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(162/255,205/255,90/255),rgb(0,0,0),rgb(85/255,107/255,47/255)), cex=3)
  dev.off()
}

```

### Visits A & B: case vs. control V Nic Alert 0-1 vs. 2-5

```{r visitABnicAlert}

# set up Nic Alert binary: 0-1 vs. 2-5
tDesignABRBE$NicAlertBinary<-as.character(tDesignABRBE$NicAlert.at.Visit)
tDesignABRBE$NicAlertBinary[which(tDesignABRBE$NicAlert.at.Visit %in% c("Level 0 (0-10 ng/mL)","Level 1 (10-30 ng/mL)"))]<-"0-1"
tDesignABRBE$NicAlertBinary[which(tDesignABRBE$NicAlert.at.Visit %in% c("Level 2 (30-100 ng/mL)","Level 3 (100-200 ng/mL)","Level 4 (200-500 ng/mL)","Level 5 (500-1000 ng/mL)"))]<-"2-5"
table(tDesignABRBE$NicAlertBinary)
tDesignABRBE$NicAlertBinary<-factor(tDesignABRBE$NicAlertBinary, levels=c("0-1","2-5"))

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$NicAlertBinary)

# visit A: case vs. control as well as Nic Alert
for (i in 1:nrow(tGSmeansABRBEV))
{
  curMod<-rownames(tGSmeansABRBEV)[i]
  curFN<-paste(curMod, ".png", sep="")
  
  statMean<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$NicAlertBinary), mean)  
  statSD<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$NicAlertBinary), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$NicAlertBinary), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_NicAlert_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(210/255,105/255,30/255),rgb(0,0,0),rgb(139/255,69/255,19/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.Nic0-1","Case.Nic0-1","Control.Nic2-5","Case.Nic2-5"), cex.axis=0.8)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(210/255,105/255,30/255, alpha=0.5), border=rgb(210/255,105/255,30/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(139/255,69/255,19/255,alpha=0.5), border=rgb(139/255,69/255,19/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(210/255,105/255,30/255),rgb(0,0,0),rgb(139/255,69/255,19/255)), cex=3)
  dev.off()
}

```

### Visits A & B: case vs. control V male vs. female

```{r visitABcaVcoVposVneg}

table(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Sex)

for (i in 1:nrow(tGSmeansABRBEV))
{
  curMod<-rownames(tGSmeansABRBEV)[i]
  curFN<-paste(curMod, ".png", sep="")
  
  statMean<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Sex), mean)  
  statSD<-tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Sex), sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBEV[i,], list(tDesignABRBE$Case.or.Control.Status.Original, tDesignABRBE$Sex), length))
  upLim<-as.vector(2^(statMean + 1.96*statSE))
  lowLim<-as.vector(2^(statMean - 1.96*statSE))
    
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseVcontrol_Gender_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(255/255,64/255,64/255),rgb(0,0,0),rgb(139/255,35/255,35/255)), cex=3, xlim=c(0.7,4.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:4, labels=c("Control.F","Case.F","Control.M","Case.M"), cex.axis=0.8)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(84/255,84/255,84/255,alpha=0.5), border=rgb(84/255,84/255,84/255,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(255/255,64/255,64/255, alpha=0.5), border=rgb(255/255,64/255,64/255, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=3.98, xright=4.02, ybottom=lowLim[4], ytop=upLim[4], col=rgb(139/255,35/255,35/255,alpha=0.5), border=rgb(139/255,35/255,35/255,alpha=0.5))
  
  points(x=1:4, y=as.vector(2^statMean), pch=19, col=c(rgb(84/255,84/255,84/255),rgb(255/255,64/255,64/255),rgb(0,0,0),rgb(139/255,35/255,35/255)), cex=3)
  dev.off()
}

```

### Visits A & B case fast vs. case slow vs. control

```{r rapidCaseAB}

tDesignABRBE$CaseControlTrip<-as.character(tDesignABRBE$Case.or.Control.Status.Original)
tDesignABRBE$CaseControlTrip[which(tDesignABRBE$Case.or.Control.Status.Original=="Case" & tDesignABRBE$Exacerbate.Within.2.Days.of.Cold=="Yes")]<-"RapidCase"
tDesignABRBE$CaseControlTrip[which(tDesignABRBE$Case.or.Control.Status.Original=="Case" & tDesignABRBE$Exacerbate.Within.2.Days.of.Cold=="No")]<-"SlowCase"
table(tDesignABRBE$CaseControlTrip)
tDesignABRBE$CaseControlTrip<-factor(tDesignABRBE$CaseControlTrip, levels=c("Control","SlowCase","RapidCase"))

table(tDesignABRBE$CaseControlTrip)

for (i in 1:nrow(tGSmeansABRBEV))
{
  curMod<-rownames(tGSmeansABRBEV)[i]
  curFN<-paste(curMod, ".png", sep="")
  
  statMean<-tapply(tGSmeansABRBEV[i,], tDesignABRBE$CaseControlTrip, mean)  
  statSD<-tapply(tGSmeansABRBEV[i,], tDesignABRBE$CaseControlTrip, sd)
  statSE<-statSD/sqrt(tapply(tGSmeansABRBEV[i,], tDesignABRBE$CaseControlTrip, length))
  upLim<-2^(statMean + 1.96*statSE)
  lowLim<-2^(statMean - 1.96*statSE)
  
  png(file=paste(file.path(resultsDir, "tempus_boxplots_meanSEplots_moduleLevel_visitAB_caseSplitBySpeedVcontrol_"), curFN, sep=""), units="in", width=3, height=4, res=300, pointsize=8)
  plot(x=1:3, y=as.vector(2^statMean), pch=19, col=c(rgb(0,0,0), rgb(0,1,0),rgb(0/255,139/255,0/255)), cex=3, xlim=c(0.7,3.3), ylim=c(0, max(upLim)), xlab="", axes=FALSE, ylab="Normalized Gene Expression", main=modNames[i])
  box()
  axis(1, at=1:3, labels=c("Control","SlowCase","RapidCase"), cex.axis=1)
  axis(2, las=2)
  # need to add segments for error bars
  rect(xleft=0.98, xright=1.02, ybottom=lowLim[1], ytop=upLim[1], col=rgb(0,0,0,alpha=0.5), border=rgb(0,0,0,alpha=0.5))
  rect(xleft=1.98, xright=2.02, ybottom=lowLim[2], ytop=upLim[2], col=rgb(0,1,0, alpha=0.5), border=rgb(0,1,0, alpha=0.5))
  rect(xleft=2.98, xright=3.02, ybottom=lowLim[3], ytop=upLim[3], col=rgb(0/255,139/255,0/255,alpha=0.5), border=rgb(0/255,139/255,0/255,alpha=0.5))
  
  points(x=1:3, y=as.vector(2^statMean), pch=19, col=c(rgb(0,0,0), rgb(0,1,0),rgb(0/255,139/255,0/255)), cex=3)
  dev.off()
}

```